### 2017-01-06 paper/references.tex
tex files tmp

\begin{thebibliography}{100}

	\bibitem{hughes_arrows_98} Generalising Monads to Arrows
	\url{https://jcp.org/en/jsr/detail?id=220}, November 10, 1998
	%
\end{thebibliography}