### 2017-01-07 paper/skeletons.tex
wip on the paper

\section{Skeletons}
With the \code{ArrowParallel} typeclass in place and implemented, we can now implement some basic parallel skeletons.

\subsection{parEvalNLazy}
\code{parEvalN} is 100\% strict, which means that it fully evaluates all passed arrows. Sometimes this might not be feasible, as it will not work on infinite lists of functions like e.g. \code{map (arr . (+)) [1..]} or just because we need the arrows evaluated in chunks. \code{parEvalNLazy} fixes this. It works by first chunking the input from \code{[a]} to \code{[[a]]} with the given \code{ChunkSize} in \code{(arr \$ chunksOf chunkSize)}. These chunks are then fed into a list \code{[arr [a] [b]]} of parallel arrows created by feeding chunks of the passed \code{ChunkSize} into the regular parEvalN by using \code{listApp}. The resulting \code{[[b]]} is lastly converted into \code{[b]} with \code{arr concat}.
\begin{lstlisting}[frame=htrbl]
parEvalNLazy :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> [arr a b] -> ChunkSize -> (arr [a] [b])
parEvalNLazy conf fs chunkSize =
	(arr $ chunksOf chunkSize) >>>
	listApp fchunks >>>
	arr concat
	where
		fchunks = map (\x -> parEvalN conf x) $ chunksOf chunkSize fs
\end{lstlisting}

\subsection{parEval2}

-- lift the functions to "maybe evaluated" functions
-- so that if they are passed a Nothing they don't compute anything
-- then, make a list of two of these functions evaluated after each other,
-- feed each function the real value and one Nothing for the function they don't have to compute
-- and combine them back to a tuple

\begin{lstlisting}[frame=htrbl]
arrMaybe :: (ArrowApply arr) => (arr a b) -> arr (Maybe a) (Maybe b)
arrMaybe fn = (arr $ go) >>> app
	where 
		go Nothing = (arr $ \Nothing -> Nothing, Nothing)
		go (Just a) = ((arr $ \(Just x) -> (fn, x)) >>> app >>> arr Just, (Just a))
\end{lstlisting}

\begin{lstlisting}[frame=htrbl]
-- evaluate two functions with different types in parallel
parEval2 :: (ArrowParallel arr a b conf,
	ArrowParallel arr (Maybe a, Maybe c) (Maybe b, Maybe d) conf,
	ArrowApply arr) =>
	conf -> arr a b -> arr c d -> (arr (a, c) (b, d))
parEval2 conf f g =
	(arr $ \(a, c) -> (f_g, [(Just a, Nothing), (Nothing, Just c)])) >>>
	app >>>
	(arr $ \comb -> (fromJust (fst (comb !! 0)), fromJust (snd (comb !! 1))))
where
	f_g = parEvalN conf $ replicate 2 $ arrMaybe f *** arrMaybe g
\end{lstlisting}

\subsection{parMap}

\begin{lstlisting}[frame=htrbl]
parMap :: (ArrowParallel arr a b conf, ArrowApply arr) =>
	conf -> (arr a b) -> (arr [a] [b])
parMap conf f = (arr $ \as -> (parEvalN conf (repeat f), as)) >>> app
\end{lstlisting}

\subsection{parMapStream}

\begin{lstlisting}[frame=htrbl]
parMapStream :: (ArrowParallel arr a b conf, ArrowChoice arr, ArrowApply arr) =>
	conf -> arr a b -> ChunkSize -> arr [a] [b]
parMapStream conf f chunkSize =
	(arr $ \as -> (parEvalNLazy conf (repeat f) chunkSize, as)) >>> app
\end{lstlisting}

\subsection{farm}

\begin{lstlisting}[frame=htrbl]
farm :: (ArrowParallel arr a b conf, ArrowParallel arr [a] [b] conf,
	ArrowChoice arr, ArrowApply arr) =>
	conf -> arr a b -> NumCores -> arr [a] [b]
farm conf f numCores =  -- chunk the input and create a function that
	-- inside of a chunk, behaves sequentially, transforms the map-chunks into a parallel function
	(arr $ \as -> (parEvalN conf $ repeat $ mapArr f, unshuffle numCores as)) >>>
	-- and then apply that function to the chunked input
	app >>>
	-- [[b]] --> [b]
	(arr shuffle)
\end{lstlisting}

\subsection{farmChunk}