### 2017-01-07 paper/benchmarks.tex
paper wip

\section{Benchmarks}
\fixme{needed for this version of the paper? we have to find better benchmarks anyways}