### 2017-01-12 paper/references.tex
added benchmarks to the paper

\begin{thebibliography}{100}

	\bibitem{hughes_arrows_98} Generalising Monads to Arrows
	\url{https://jcp.org/en/jsr/detail?id=220}, November 10, 1998

	\bibitem{marlow_parconc} Sample code for Simon Marlow's book "Parallel and Concurrent Programming in Haskell"
	\url{https://github.com/simonmar/parconc-examples}, retrieved on 01/12/2017
\end{thebibliography}