### 2017-01-12 paper/syntacticSugar.tex

\section{Syntactic Sugar}
To make using our new API a bit easier, we also introduce some syntactic sugar.
We start with \code{|>>>|}, which is basically \code{>>>} on lists of arrows.
\begin{lstlisting}[frame=htrbl]
(|>>>|) :: (Arrow arr) => [arr a b] -> [arr b c] -> [arr a c]
(|>>>|) = zipWith (>>>)
\end{lstlisting}
For basic arrows, we have the \code{***} combinator which allows us to combine two arrows \code{arr a b} and \code{arr c d} into an arrow \code{arr (a, c) (b, d)} which does both computations at once. This can easily be translated into a parallel version with \code{parEval2}, but for this we require a backend which has an implementation that does not require any configuration (hence the \code{()} as the conf parameter in the following code snippet).
\begin{lstlisting}[frame=htrbl]
(|***|) :: (ArrowParallel arr a b (),
	ArrowParallel arr (Maybe a, Maybe c) (Maybe b, Maybe d) (),
	ArrowApply arr) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{lstlisting}
With this we can analogously to the serial \code{&&&} define the parallel \code{|&&&|}.
\begin{lstlisting}[frame=htrbl]
(|&&&|) :: (ArrowParallel arr a b (),
	ArrowParallel arr (Maybe a, Maybe a) (Maybe b, Maybe c) (),
	ArrowApply arr) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{lstlisting}