### 2017-01-12 paper/parallelHaskells.tex

\section{Short introduction to parallel Haskells}
There are already several ways write parallel programs in Haskell. As we will base our parallel arrows on existing parallel Haskells, we will now give a short introduction to the ones we use as backends in this paper.

\subsection{Multicore Haskell}

\begin{lstlisting}[frame=htrbl]
parallel :: (NFData b) => [a -> b] -> [a] -> [b]
parallel fs as = zipWith ($) fs as `using` parList rdeepseq
\end{lstlisting}

\subsection{ParMonad}

\begin{lstlisting}[frame=htrbl]
parallel :: (NFData b) => [a -> b] -> [a] -> [b]
parallel fs as = runPar $ 
	(sequenceA $ map (spawnP) $ zipWith ($) fs as) >>= mapM get
\end{lstlisting}

\subsection{Eden}

\begin{lstlisting}[frame=htrbl]
parallel :: (Trans a, Trans b) => [a -> b] -> [a] -> [b]
parallel fs as = spawnF fs as
\end{lstlisting}