### 2017-01-12 paper/motivation.tex
wip on the paper

\section{Motivation}
Arrows were introduced in John Hughes paper as an alternative to monads for API design \citHughes. In the paper Hughes describes that arrows have one powerful additional property when compared to monads for API design: Extensibility. In this paper we will show how this property can be used to add parallelism capabilities to different arrow-types.