### 2017-01-13 paper/references.tex

\begin{thebibliography}{100}

	\bibitem{hughes_arrows_98} Hughes, Generalising Monads to Arrows
	\url{https://jcp.org/en/jsr/detail?id=220}, November 10, 1998

	\bibitem{marlow_parconc} Sample code for Simon Marlow's book "Parallel and Concurrent Programming in Haskell"
	\url{https://github.com/simonmar/parconc-examples}, retrieved on 01/12/2017

	\bibitem{multicore_hackage} Multicore's parallel package on Hackage
	\url{https://hackage.haskell.org/package/parallel-3.2.1.0}, retrieved on 01/12/2017

	\bibitem{monad_par_paper} Marlow et al, A Monad for Deterministic Parallelism
	\url{http://www.cs.indiana.edu/~rrnewton/papers/haskell2011_monad-par.pdf}, retrieved on 01/12/2017

	\bibitem{monad_par_hackage} ParMonad (Control.Monad.Par) on Hackage
	\url{https://hackage.haskell.org/package/monad-par-0.3.4.8/}, retrieved on 01/12/2017

	\bibitem{eden_hackage} Eden Modules on Hackage
	\url{https://hackage.haskell.org/package/edenmodules-1.2.0.0/}, retrieved on 01/12/2017

	\bibitem{programming_with_arrows} Hughes, Programming with Arrows
	\url{http://www.cse.chalmers.se/~rjmh/afp-arrows.pdf}, retrieved on 02/12/2017

	\bibitem{eden_skel_shuffle} Eden Skeletons' Control.Parallel.Eden.Map package source code
	\url{https://hackage.haskell.org/package/edenskel-2.1.0.0/docs/src/Control-Parallel-Eden-Map.html}, retrieved on 02/12/2017

\end{thebibliography}