### 2017-01-19 paper/parrows.tex

\section{Parallel Arrows}
We have seen what Arrows are and how they can be used as a general interface to computation. In the following section we will discuss how Arrows can also be looked at as a general interface not only to computation, but to \textbf{parallel computation} as well. We start by introducing the interface and explaining the reasonings behind it. Then, we discuss some implementations using exisiting parallel Haskells. Finally, we explain why using Arrows for expressing parallelism is beneficial.
\subsection{The ArrowParallel typeclass}
As we have seen earlier, in its purest form, parallel computation (on functions) can be looked at as the execution of some functions \code{a -> b} in parallel:
\begin{lstlisting}[frame=htrbl]
parEvalN :: [a -> b] -> [a] -> [b]
\end{lstlisting}
Translating this into arrow terms gives us a new operator \code{parEvalN} that lifts a list of arrows \code{[arr a b]} to a parallel arrow \code{arr [a] [b]} (This combinator is similar to our utility function listApp, but does parallel instead of serial evaluation).
\begin{lstlisting}[frame=htrbl]
parEvalN :: (Arrow arr) => [arr a b] -> arr [a] [b]
\end{lstlisting}
With this definition of \code{parEvalN}, parallel execution can be looked at as yet another arrow combinator. But as the implementation may differ depending on the actual type of the arrow \code{arr} and we want this to be an interface for different backends, we have to introduce the new typeclass \code{ArrowParallel} to host this combinator.
\begin{lstlisting}[frame=htrbl]
class Arrow arr => ArrowParallel arr a b where
	parEvalN :: [arr a b] -> arr [a] [b]
\end{lstlisting}
Sometimes parallel Haskells require additional configuration parameters as information about the execution environment. This is why we also introduce an additional \code{conf} parameter to the function. We also do not want \code{conf} to be a fixed type, as the configuration parameters can differ for different instances of \code{ArrowParallel}. So we add it to the type signature of the typeclass as well.
\begin{lstlisting}[frame=htrbl]
class Arrow arr => ArrowParallel arr a b conf where
	parEvalN :: conf -> [arr a b] -> arr [a] [b]
\end{lstlisting}
We don't require the conf parameter in every implementation. If it is not needed, we usually want to allow the \code{conf} type parameter to be of any type and don't even evaluate it by blanking it in the type signature of the implemented \code{parEvalN} as we will see in the implementation of the Multicore and the ParMonad backend.

\subsection{Multicore Haskell}
The Multicore Haskell implementation of this class is straightforward using listApp from chapter \ref{utilfns} combined with the \code{using} operator from Multicore Haskell.
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b conf where
		parEvalN _ fs = listApp fs >>> arr (flip using $ parList rdeepseq)
\end{lstlisting}
We hardcode the \code{parList rdeepseq} strategy here, as in this context it is the only one making sense, since we usually want the output list to be fully evaluated to its normal form.

\subsection{ParMonad}
The ParMonad implementation makes use of Haskells laziness and ParMonad's \code{spawnP :: NFData a => a -> Par (IVar a)} function, which forks away the computation of a value and returns an IVar containing the result in the Par monad.
\\\\
We therefore apply each function to its corresponding input value with \code{app} and then fork the computation away with \code{arr spawnP} inside a \code{zipWithArr} call. This yields a list \code{[Par (IVar b)]}, which we then convert into \code{Par [IVar b]} with \code{arr sequenceA}. In order to wait for the computation to finish, we map over the \code{IVar}s inside the ParMonad with \code{arr (>>= mapM get)}. The result of this operation is a \code{Par [b]} from which we can finally remove the monad again by running \code{arr runPar} to get our output of \code{[b]}.
\begin{lstlisting}[frame=htrbl]
instance (NFData b, ArrowApply arr, ArrowChoice arr) =>
	ArrowParallel arr a b conf where
		parEvalN _ fs = 
			(arr $ \as -> (fs, as)) >>>
			zipWithArr (app >>> arr spawnP) >>>
			arr sequenceA >>>
			arr (>>= mapM get) >>>
			arr runPar
\end{lstlisting}

\subsection{Eden}
For the Multicore and ParMonad implementation we could use general instances of \code{ArrowParallel} that just require the \code{ArrowApply} and \code{ArrowChoice} typeclasses. With Eden this is not the case as we can only spawn a list of functions and we cannot extract simple functions out of arrows. While we could still manage to have only one class in the module by introducing a typeclass like
\begin{lstlisting}[frame=htrbl]
class (Arrow arr) => ArrowUnwrap arr where
	arr a b -> (a -> b)
\end{lstlisting}
, we don't do it in this paper, as this seems too hacky. For now, we just implement \code{ArrowParallel} for normal functions
\begin{lstlisting}[frame=htrbl]
instance (Trans a, Trans b) => ArrowParallel (->) a b conf where
parEvalN _ fs as = spawnF fs as
\end{lstlisting}
and the Kleisli type.
\begin{lstlisting}[frame=htrbl]
instance (Monad m, Trans a, Trans b, Trans (m b)) =>
	ArrowParallel (Kleisli m) a b conf where
parEvalN conf fs =
	(arr $ parEvalN conf (map (\(Kleisli f) -> f) fs)) >>>
	(Kleisli $ sequence)
\end{lstlisting}

\subsection{Benefits of parallel Arrows}
We have seen that we can wrap parallel Haskells inside of the \code{ArrowParallel} interface, but why do we abstract parallelism this way and what does this approach do better than the other parallel Haskells?
\begin{itemize}
	\item \textbf{Arrow API benefits}:
	With the \code{ArrowParallel} typeclass we do not lose any benefits of using arrows as \code{parEvalN} is just yet another arrow combinator. The resulting arrow can be used in the same way its serial version can be used. This is a big advantage of this approach, especially compared to the monad solutions as all arrow combinators will continue to work instead of requiring special ones for each parallel monad.
	\item \textbf{Abstraction}:
	With the \code{ArrowParallel} typeclass, we abstracted all parallel implementation logic away from the business logic. This gives us the beautiful situation of being able to write our code against the interface the typeclass gives us without being bound to any parallel Haskell. So as an example, during development, we can run the code on the simple Multicore version and afterwards deploy it on a cluster using an Eden built, by just swapping out the actual \code{ArrowParallel} instance.
\end{itemize}