### 2017-01-26 presentation/benchmarks.tex
copied all the relevant content from the paper to the presentation

\section{Benchmarks}
The Benchmarks were run on a Core i7-3970X CPU @ 3.5GHz with 6 cores and 12 threads. For sake of comparability with Simon Marlow's parallel version which uses the ParMonad, we use the ParMonad backend for the parallel arrow versions as well.
\frbreak
\framebox{
	\begin{tikzpicture}[thick, scale=0.7]
	\begin{axis}[
	xbar, xmin=0,
	width=12cm, height=7cm, enlarge y limits=0.5,
	xlabel={avg. seconds run on sudoku17.1000.txt (100 runs)},
	symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
	ytick=data,
	nodes near coords, nodes near coords align={horizontal},
	]
	\addplot coordinates {(.215,parrows-sudoku-farmChunk) (.213,parrows-sudoku-farm) (.188,parrows-sudoku-parmapstream) (.197,parrows-sudoku-parmap) (.527,parmonad-sudoku-parmap) (1.338,sudoku-sequential)};
	\end{axis}
	\end{tikzpicture}
}
\frbreak
\framebox{
	\begin{tikzpicture}[thick, scale=0.7]
	\begin{axis}[
	xbar, xmin=0,
	width=12cm, height=7cm, enlarge y limits=0.5,
	xlabel={avg. seconds run on sudoku17.16000.txt (100 runs)},
	symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
	ytick=data,
	nodes near coords, nodes near coords align={horizontal},
	]
	\addplot coordinates {(3.218,parrows-sudoku-farmChunk) (3.148,parrows-sudoku-farm) (2.755,parrows-sudoku-parmapstream) (2.860,parrows-sudoku-parmap) (2.936,parmonad-sudoku-parmap) (21.324,sudoku-sequential)};
	\end{axis}
	\end{tikzpicture}
}
\frbreak
\framebox{
	\begin{tikzpicture}[thick, scale=0.7]
	\begin{axis}[
	xbar, xmin=0,
	width=12cm, height=7cm, enlarge y limits=0.5,
	xlabel={avg. seconds run on sudoku17.49151.txt (100 runs)},
	symbolic y coords={parrows-sudoku-farmChunk, parrows-sudoku-farm, parrows-sudoku-parmapstream, parrows-sudoku-parmap, parmonad-sudoku-parmap, sudoku-sequential},
	ytick=data,
	nodes near coords, nodes near coords align={horizontal},
	]
	\addplot coordinates {(10.038,parrows-sudoku-farmChunk) (10.151,parrows-sudoku-farm) (8.776,parrows-sudoku-parmapstream) (8.969,parrows-sudoku-parmap) (8.679,parmonad-sudoku-parmap) (60.931,sudoku-sequential)};
	\end{axis}
	\end{tikzpicture}
}