### 2017-01-31 presentation/skeletons.tex
wip presentation

\section{Usability}
\subsection{Skeletons}
\begin{frame}[fragile]{Skeletons... (1)}
parEvalN, but \textbf{chunky}:
\begin{lstlisting}[frame=htrbl]
parEvalNLazy :: conf -> ChunkSize -> [arr a b] -> (arr [a] [b])
\end{lstlisting}
parallel evaluation of \textbf{different typed functions}:
\begin{lstlisting}[frame=htrbl]
parEval2 :: conf -> arr a b -> arr c d -> (arr (a, c) (b, d))
\end{lstlisting}
\end{frame}
\begin{frame}[fragile]{Skeletons... (2)}
map, but in parallel:
\begin{lstlisting}[frame=htrbl]
parMap :: conf -> (arr a b) -> (arr [a] [b])
\end{lstlisting}
parMap, but \textbf{chunky}:
\begin{lstlisting}[frame=htrbl]
parMapStream :: conf -> ChunkSize -> arr a b -> arr [a] [b]
\end{lstlisting}
\end{frame}
\begin{frame}[fragile]{Skeletons... (3)}
parMap, but with \textbf{workload distribution}:
\begin{lstlisting}[frame=htrbl]
farm :: conf -> NumCores -> arr a b -> arr [a] [b]
\end{lstlisting}
farm, but \textbf{chunky}:
\begin{lstlisting}[frame=htrbl]
farmChunk ::
	conf -> ChunkSize -> NumCores -> arr a b-> arr [a] [b]
\end{lstlisting}
\end{frame}