### 2017-02-20 presentation/utilityFunctions.tex

\subsection{Utility Functions}\label{utilfns}
\begin{frame}[fragile]{mapArr, zipWithArr, listApp}
From \cite{programming_with_arrows}, we have:
\begin{lstlisting}[frame=htrbl]
mapArr :: ArrowChoice arr => arr a b -> arr [a] [b]
\end{lstlisting}

zipWithArr lifts any arrow \code{arr (a, b) c} to an arrow \code{arr ([a], [b]) [c]}:
\begin{lstlisting}[frame=htrbl]
zipWithArr :: ArrowChoice arr => arr (a, b) c -> arr ([a], [b]) [c]
zipWithArr f = (arr $ \(as, bs) -> zipWith (,) as bs) >>>
	mapArr f
\end{lstlisting}
listApp converts a list of arrows \code{[arr a b]} to a new arrow \code{arr [a] [b]}:
\begin{lstlisting}[frame=htrbl]
listApp :: (ArrowChoice arr, ArrowApply arr) =>
	[arr a b] -> arr [a] [b]
listApp fs = (arr $ \as -> (fs, as)) >>> zipWithArr app
\end{lstlisting}
with the \code{ArrowApply} that defines \code{app :: arr (arr a b, a) c}.
\end{frame}