### 2017-03-30 paper/abstract.tex
merged with Oleg's first changes. only topologySkeletons.tex has to be merged

\begin{abstract}
Arrows form a general interface for computation and pose therefore as an alternative to monads for API design. In the paper Hughes describes how arrows are a generalization of monads and how they are not as restrictive. We express parallelism using this concept. We define an Arrows-based interface for parallel computations and implement it using multiple parallel Haskells.

This new way of writing parallel programs has a benefit of being portable across flavours of parallel Haskells used. With our framework we define  several parallel skeletons. We also  introduce some syntactic sugar to provide parallelism-aware arrow combinators similar to the traditional ones. Benchmarks shows that our framework does not induce too much overhead performance-wise.
\end{abstract}