### 2017-05-11 paper/abstract.tex
applied olegs patches

\begin{abstract}
Arrows form\olcomment{are?} a general interface for computation and pose therefore as an alternative to monads for API design. We express parallelism using this concept. This is a new way to represent parallel computation. We define an Arrows-based interface for parallelism and implement it using multiple parallel Haskells.
\olcomment{Benefits:} In this manner we are able to bridge across various parallel Haskells with a common interface.

This new way of writing parallel programs has a benefit of being portable across flavours of parallel Haskells.\olcomment{Wdh?}
Each parallel computation is an arrow, they can be composed and transformed as such.
We thus introduce some syntactic sugar to provide parallelism-aware arrow combinators.

We also define  several parallel skeletons with our framework. 
Benchmarks shows that our framework does not induce too much overhead performance-wise.
\end{abstract}