### 2017-05-26 paper/content/syntacticSugar.tex
wip

\subsubsection{Syntactic Sugar} \label{syntacticSugar}
For basic arrows, we have the |***| combinator (Fig.~\ref{fig:syntacticSugarArrows}) which allows us to combine two arrows |arr a b| and |arr c d| into an arrow |arr (a, c) (b, d)| which does both computations at once. This can easily be translated into a parallel version |parstar| (Fig.~\ref{fig:par***}) with the use of |parEval2|, but for this we require a backend which has an implementation that does not require any configuration (hence the |()| as the |conf| parameter% in Fig.~\ref{fig:par***}
):
% \begin{figure}[h]
\begin{code}
(|***|) :: (ArrowChoice arr, ArrowParallel arr (Either a c) (Either b d) ())) =>
	arr a b -> arr c d -> arr (a, c) (b, d)
(|***|) = parEval2 ()
\end{code}
% \caption{Definition of |parstar|, the parallel version of |***|.}
% \label{fig:par***}
% \end{figure}
% With this we can analogously to the serial |&&&|
We define the parallel |parand| % (Fig.~\ref{fig:par&&&}) 
in a similar manner to its sequential pendant |&&&| (Fig.~\ref{fig:syntacticSugarArrows}):
% \begin{figure}[h]
\begin{code}
(|&&&|) :: (ArrowChoice arr, ArrowParallel arr (Either a a) (Either b c) ()) =>
	arr a b -> arr a c -> arr a (b, c)
(|&&&|) f g = (arr $ \a -> (a, a)) >>> f |***| g
\end{code} % $ %% formatting
% \caption{Definition of |parand| - the parallel version of |&&&|.}
% \label{fig:par&&&}
% \end{figure}