### 2017-07-03 paper/content/mapSkeletons.tex
added olegl changes

%%% \FloatBarrier
\section{Map-based Skeletons}
\label{sec:map-skeletons}
Now we have developed Parallel Arrows far enough to define some useful algorithmic skeletons that abstract typical parallel computations.

\paragraph{Parallel |map| and laziness.}
The |parMap| skeleton (Figs.~\ref{fig:parMapImg},~\ref{fig:parMap}) is probably the most common skeleton for parallel programs. We can implement it with |ArrowParallel| by repeating an arrow |arr a b| and then passing it into |parEvalN| to obtain an arrow |arr [a] [b]|.
Just like |parEvalN|, |parMap| is 100\% strict.
As |parMap|  is 100\% strict it has the same restrictions as |parEvalN| compared to |parEvalNLazy|. So it makes sense to also have a |parMapStream| (Figs.~\ref{fig:parMapStreamImg},~\ref{fig:parMapStream}) which behaves like |parMap|, but uses |parEvalNLazy| instead of |parEvalN|. The code is quite straightforward, we show it in Appendix.

\begin{figure}[thb]
%farm
\includegraphics[scale=0.7]{images/farm}
\caption{Schematic depiction of a |farm|, a statically
      load-balanced |parMap|.}
\label{fig:farmImg}

\begin{code}
farm :: (ArrowParallel arr a b conf,
	ArrowParallel arr [a] [b] conf, ArrowChoice arr) =>
	conf -> NumCores -> arr a b -> arr [a] [b]
farm conf numCores f =
	unshuffle numCores >>>
	parEvalN conf (repeat (mapArr f)) >>>
	shuffle
\end{code}
\caption{The definition of |farm|.}
\label{fig:farm}

%farmChunk
\includegraphics[scale=0.7]{images/farmChunk}
\caption{Schematic depiction of |farmChunk|.}
\label{fig:farmChunkImg}
\end{figure}

\paragraph{Statically load-balancing parallel |map|.}
Our |parMap| spawns every single computation in a new thread (at least for the instances of |ArrowParallel| we gave in this paper). This can be quite wasteful and a |farm| (Figs.~\ref{fig:farmImg},~\ref{fig:farm}) that equally distributes the workload over |numCores| workers seems useful. \olcomment{it is a \emph{static} load balancing}
The definitions of the helper functions |unshuffle|, |takeEach|, |shuffle| (Fig.~\ref{fig:edenshuffleetc}) originate from an Eden skeleton\footnote{Available on Hackage under \url{https://hackage.haskell.org/package/edenskel-2.1.0.0/docs/src/Control-Parallel-Eden-Map.html}.}.

%\paragraph{Lazy statically load-balancing parallel map}
Since a |farm|  is basically just |parMap| with a different work distribution, it is, again, 100\% strict. So we can define |farmChunk| (Figs.~\ref{fig:farmChunkImg},~\ref{fig:farmChunk}) which uses |parEvalNLazy| instead of |parEvalN|. It is basically the same definition as for |farm|, with |parEvalN| replaced with |parEvalNLazy|.

%%% Local Variables:
%%% mode: latex
%%% TeX-master: "main"
%%% End: