### 2017-07-08 paper/content/benchmarks_tmp.tex
minor

\documentclass[12pt,a4paper]{article}

\usepackage{microtype} %% looks better
\usepackage{pgfplots}

\usepackage{pgfplotstable} %% for regression

\begin{document}

\newcommand{\performanceplot}[5]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
xlabel=threads,
xtick={#4},
ylabel=time (seconds),
grid=major,
legend entries={#2},
legend style={at={(0.99,0.99)},anchor=north east},
max space between ticks=50pt,
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
xmin=-1,
xmax=#3+1]
#5
\end{axis}
\end{tikzpicture}
}

\newcommand{\performancediffplot}[6]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
xlabel=threads,
xtick={#4},
ytick distance=#6,
minor tick num=9,
ylabel=time (seconds),
grid=both,
legend entries={#2},
legend style={at={(0.93,0.01)},anchor=south east},
max space between ticks=50pt,
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
xmin=-1,
xmax=#3+1]
#5
\end{axis}
\end{tikzpicture}
}

\newcommand{\speedupplot}[5]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
xlabel=threads,
xtick={#4},
ytick={#4},
ylabel=speedup,
grid=major,
legend entries={linear, #2},
legend style={at={(0.01,0.99)},anchor=north west},
max space between ticks=50pt,
enlargelimits={abs=1 * #3/32},
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
ymin=-1,
xmin=-1,
ymax=#3 + 1,
xmax=#3 + 1]
\addplot [domain=0:#3, no markers,dotted,thick]{x};
#5
\end{axis}
\end{tikzpicture}
}

\newcommand{\speedupdiffplot}[5]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
xlabel=threads,
xtick={#4},
ylabel=time (seconds),
grid=major,
legend entries={#2},
legend style={at={(0.93,0.01)},anchor=south east},
max space between ticks=50pt,
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
xmin=-1,
xmax=#3+1]
#5
\end{axis}
\end{tikzpicture}
}

\newcommand{\performanceSkelRMSM}[2]{
\performanceplot{Parallel performance of \\ SkelRM #1 #2}{Eden CP, Multicore, Par Monad}{32}{2,4,8,16,32}{
\addplot+ table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-#1-#2.csv};
\addplot+ table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-#1-#2.csv};
\addplot+ table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-par-#1-#2.csv};
}
}

\newcommand{\speedupSkelRMSM}[2]{
\speedupplot{Parallel speedup of \\ SkelRM #1 #2}{Eden CP, Multicore, Par Monad}{32}{2,4,8,16,32}{
\addplot+ table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-#1-#2.csv};
\addplot+ table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-#1-#2.csv};
\addplot+ table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/sm-rm/bench-sm-rm.bench.skelrm-parr-par-#1-#2.csv};
}
}

\performanceSkelRMSM{4423}{32}

\speedupSkelRMSM{4423}{32}

\performanceSkelRMSM{4423}{64}

\speedupSkelRMSM{4423}{64}

\performanceSkelRMSM{9941}{32}

\speedupSkelRMSM{9941}{32}

\performanceSkelRMSM{9941}{64}

\speedupSkelRMSM{9941}{64}

\performanceSkelRMSM{11213}{32}

\speedupSkelRMSM{11213}{32}

\performanceSkelRMSM{11213}{64}

\speedupSkelRMSM{11213}{64}

\newcommand{\performanceSkelRMDist}[3]{
\performanceplot{Parallel performance of \\ SkelRM #1 #2}{PArrows, Eden}{256}{#3}{
\addplot table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/distributed-rm/bench-distributed.bench.skelrm-parrows-#1-#2.csv};
\addplot table [scatter, x="nCores", y="time", col sep=comma, mark=none,
smooth]{benchmarks/distributed-rm/bench-distributed.bench.skelrm-eden-#1-#2.csv};
}
}

\newcommand{\performanceSkelRMDistDiff}[4]{
\performancediffplot{Parallel performance \\ difference of SkelRM #1 #2}{(Eden - PArrows)}{256}{#3}{
\addplot+[mark=*,thick,error bars/.cd,
    y dir=both,y explicit] table [x="nCores", y="time", y error="max stddev", col sep=comma, mark=dots,
smooth]{benchmarks/distributed-rm/#1-#2-diff.csv};
}{#4}
}

\newcommand{\speedupSkelRMDist}[3]{
\speedupplot{Parallel speedup of \\ SkelRM #1 #2}{PArrows, Eden}{256}{#3}{
\addplot table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/distributed-rm/bench-distributed.bench.skelrm-parrows-#1-#2.csv};
\addplot table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/distributed-rm/bench-distributed.bench.skelrm-eden-#1-#2.csv};
}
}

\newcommand{\speedupSkelRMDistDiff}[3]{
\speedupdiffplot{Parallel speedup \\ difference of SkelRM #1 #2}{(Eden - PArrows)}{256}{#3}{
\addplot table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{benchmarks/distributed-rm/#1-#2-diff.csv};
}
}

\performanceSkelRMDist{9941}{256}{16,32,64,128,256}

\performanceSkelRMDistDiff{9941}{256}{16,32,64,128,256}{0.11}

\speedupSkelRMDist{9941}{256}{16,32,64,128,256}

\speedupSkelRMDistDiff{9941}{256}{16,32,64,128,256}

\performanceSkelRMDist{11213}{256}{16,32,64,128,256}

\performanceSkelRMDistDiff{11213}{256}{16,32,64,128,256}{0.1}

\speedupSkelRMDist{11213}{256}{16,32,64,128,256}

\speedupSkelRMDistDiff{11213}{256}{16,32,64,128,256}

\performanceSkelRMDist{44497}{256}{32,64,128,256}

\performanceSkelRMDistDiff{44497}{256}{32,64,128,256}{0.5}

\speedupSkelRMDist{44497}{256}{32,64,128,256}

\speedupSkelRMDistDiff{44497}{256}{32,64,128,256}

\end{document}