### 2017-09-25 paper/content/lhs2TeXSafeCommands.tex
wip on parallelHaskells.tex (added overview table)

\newcommand{\parallelHaskellOverview}{
\begin{table}
\caption{Overview over some parallel Haskells}
\centering
%% something was wrong with separators in table
\renewcommand{\tabcolsep}{1em}
\begin{tabular}{ccc}
\toprule
Name & Memory setting & Concepts used \\
\midrule
GpH         & shared \& distributed & \ensuremath{par}/\ensuremath{pseq} "hints" \\
\ensuremath{Par} Monad & shared & deterministic parallel Monad \\
Eden        & shared \& distributed & special compiler, pure \& monadic interface \\
HdpH        & \mbcomment{shared?} \& distributed & parallel Monad, Template Haskell \\
LVish       & shared & parallel Monad, parallel data-structures \\
\bottomrule
\end{tabular}
\end{table}
}