### 2017-10-17 paper/content/futures.tex
minor fixes

\section{Futures} \label{sec:futures}
Consider a mock-up parallel arrow combinator:
\begin{code}
someCombinator :: (Arrow arr) => [arr a b] -> [arr b c] -> arr [a] [c]
someCombinator fs1 fs2 = parEvalN () fs1 >>> rightRotate >>> parEvalN () fs2
\end{code}

In a distributed environment, a resulting arrow of this combinator first evaluates all |[arr a b]| in parallel, sends the results back to the master node, rotates the input once and then evaluates the |[arr b c]| in parallel to then gather the input once again on the master node.
Such situations arise, \eg in scientific computations when the data distributed across the nodes needs to be transposed. A concrete example is 2D FFT computation \cite{Gorlatch,Berthold2009-fft}.

While the above example could be rewritten into only one |parEvalN| call by directly wiring the arrows together before spawning, it illustrates an important problem. When using a |ArrowParallel| backend that resides on multiple computers, all communication between the nodes is done via the master node, as shown in the Eden trace in Figure~\ref{fig:withoutFutures}. This can become a serious bottleneck %in heavy threaded applications.
for larger amount of data and number of processes \citep[as e.g.][showcases]{Berthold2009-fft}.
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.9\textwidth]{images/withoutFutures}
	\caption[without Futures]{Communication between 4 Eden processes without Futures. All communication goes through the master node. Each bar represents one process. Black lines between processes represent communication. Colors: blue $\hat{=}$ idle, green $\hat{=}$ running, red  $\hat{=}$ blocked, yellow $\hat{=}$ suspended.}
	\label{fig:withoutFutures}
\olcomment{more practical and heavy-weight example! fft (I have the code)?}\\
\mbcomment{Depends... Are the communications easy to read in such an example?}\\
\mbcomment{Keep the description for the different colours, or link to
  the EdenTV description in \ref{sec:edentv}}
\olcomment{ok as is}
\olcomment{use the fft example (when it works)?}
\end{figure}

This is in fact only a problem in distributed memory (in the scope of this paper) and we should  allow nodes to communicate directly with each other. Eden already ships with \enquote{remote data} that enable this \cite{AlGo03a,Dieterle2010}.
But as we want code with our DSL to be implementation agnostic, we have to wrap this context. We do this with the |Future| typeclass (Fig.~\ref{fig:future}).
\begin{figure}[h]
\begin{code}
class Future fut a | a -> fut where
    put :: (Arrow arr) => arr a (fut a)
    get :: (Arrow arr) => arr (fut a) a
\end{code}
\caption{Definition of the |Future| typeclass.}
\label{fig:future}
\end{figure}
Since |RD| is only a type synonym for a communication type that Eden uses internally, we have to use some wrapper classes to fit that definition, though, as Fig.~\ref{fig:RDFuture} shows. %This is due to the same reason we had to introduce a wrapper for |Strategy a| in the GpH Haskell implementation of |ArrowParallel| in Section~\ref{sec:parrows:multicore}.
Technical details are in Appendix, in Section~\ref{app:omitted}.

For the |Par| Monad and GpH Haskell backends, we can simply use |BasicFuture|s (Fig.~\ref{fig:BasicFuture}), which are just simple wrappers around the actual data with boiler-plate logic so that the typeclass is satisfied. This is because the concept of a |Future| does not change anything for shared-memory execution as there are no communication problems to fix. Nevertheless, we require a common interface so the parallel Arrows are portable across backends.
\begin{figure}[tb]
\begin{code}
data BasicFuture a = BF a

instance (NFData a) => NFData (BasicFuture a) where
    rnf (BF a) = rnf a

instance (NFData a) => Future BasicFuture a where
    put = arr BF
    get = arr (\(BF a) -> a)
\end{code}
\caption{The |BasicFuture| type and its |Future| instance for the |Par| Monad and GpH Haskell backends.}
\label{fig:BasicFuture}
\end{figure} % $

In our communication example we can use this |Future| concept for direct communication between nodes as shown in Fig.~\ref{fig:someCombinatorParallel}.
\begin{figure}[tbh]
\begin{code}
someCombinator :: (Arrow arr) => [arr a b] -> [arr b c] -> arr [a] [c]
someCombinator fs1 fs2 =
	parEvalN () (map (>>> put) fs1) >>>
	rightRotate >>>
	parEvalN () (map (get >>>) fs2)
\end{code}
\caption{The mock-up combinator in parallel.}
\label{fig:someCombinatorParallel}
\end{figure}
In a distributed environment, this gives us a communication scheme with messages going through the master node only if it is needed -- similar to what is shown in the trace visualization in Fig.~\ref{fig:withFutures}. \olcomment{Fig. is not really clear. Do Figs with a lot of load? --- fft?}
\begin{figure}[ht]
	\centering
	\includegraphics[width=0.9\textwidth]{images/withFutures}
	\caption[with Futures]{Communication between 4 Eden processes with Futures. Other than in Fig.~\ref{fig:withoutFutures}, processes communicate directly (black lines between the bars, one example message is marked in the Figure) instead of always going through the master node (bottom bar).}
	\label{fig:withFutures}
\end{figure}