### 2017-11-24 paper/content/main.tex

% This is our submission, modified from:
% the file JFP2egui.lhs
% release v1.02, 27th September 2001
%   (based on JFPguide.lhs v1.11 for LaLhs 2.09)
% Copyright (C) 2001 Cambridge University Press

\NeedsTeXFormat{LaTeX2e}

\documentclass{jfp1}

%%% Macros for the guide only %%%
%\providecommand\AMSLaTeX{AMS\,\LaTeX}
%\newcommand\eg{\emph{e.g.}\ }
%\newcommand\etc{\emph{etc.}}
%\newcommand\bcmdtab{\noindent\bgroup\tabcolsep=0pt%
%  \begin{tabular}{@{}p{10pc}@{}p{20pc}@{}}}
%\newcommand\ecmdtab{\end{tabular}\egroup}
%\newcommand\rch[1]{$\longrightarrow\rlap{$#1$}$\hspace{1em}}
%\newcommand\lra{\ensuremath{\quad\longrightarrow\quad}}

\jdate{August 2017}
\pubyear{2017}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\doi{...}

%\newtheorem{lemma}{Lemma}[section]

\input{lhs2TeXSafeCommands}
\input{prelude}

\title{Arrows for Parallel Computations}
\ifthenelse{\boolean{anonymous}}{%
\author{Submission ID xxxxxx}
}{%
%\author{Martin Braun, Phil Trinder, and Oleg Lobachev}
%\affiliation{University Bayreuth, Germany and Glasgow University, UK}
 \author[M. Braun, O. Lobachev and P. Trinder]%
        {MARTIN BRAUN\\
         University Bayreuth, 95440 Bayreuth, Germany\\
		 OLEG LOBACHEV\\
		 University Bayreuth, 95440 Bayreuth, Germany\\
		 \and\ PHIL TRINDER\\
		 Glasgow University, Glasgow, G12 8QQ, Scotland}
}% end ifthenelse

\begin{document}

\label{firstpage}

\def\SymbReg{\textsuperscript{\textregistered}}

\maketitle

%% environment inside
\input{abstract.tex}

\tableofcontents

	%
	%%include abstract.lhs
	%
	%\newpage
	%\pagebreak
	\input{motivation}
	\input{relwork}
	\begin{flushright}

	\end{flushright}\section{Background}
	\label{sec:background}
	\input{arrows}
	\input{parallelHaskells}
	%\pagebreak
	%\pagebreak
	%\pagebreak
	\input{parrows}
	%\pagebreak
	\input{basicSkeletons}
	%\pagebreak
	\input{syntacticSugar}
	%\pagebreak
	\input{futures}
	%\pagebreak
	\input{mapSkeletons}
	%\pagebreak
	\input{topologySkeletons}
	%\pagebreak
	\input{benchmarks}
	%%\pagebreak
	\input{conclusion}
	%\pagebreak
        \bibliographystyle{jfp}
	\bibliography{references,main}
        \appendix
	\input{utilityFunctions}
\end{document}