### 2017-11-28 paper/content/bestAndWorstBenchmarks.tex
fixed tables again

%\bestAndWorstBenchmarks

\newcommand{\meanOverheadTEST}{
\begin{table}[]
\centering
\caption{Overhead in the benchmarks}
\label{tab:meanOverhead}
\captionsetup[subtable]{justification=centering}
\subfloat[\sudokutest 1000]{
	\begin{minipage}[t]{\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{3cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{Benchmark} & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \toprule
		\csvreader[head to column names]{benchmarks/sudoku-sm/bestAndWorstSudoku-1000.csv}{}{\vs & \benchmark \params & \meanOverhead & \maxStdDevForOverhead \\}
		\midrule
	\end{tabular}
	\endgroup
	\label{sudokuMeanFactor-1000}
	\end{minipage}
}
\caption{Overhead in the benchmarks}
\end{table}
}

\newcommand{\meanOverhead}{
\begin{table}[]
\centering
\caption{Overhead in the benchmarks}
\label{tab:meanOverhead}
\captionsetup[subtable]{justification=centering}
\subfloat[\sudokutest 1000]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/sudoku-sm/bestAndWorstSudoku-1000.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
	\label{sudokuMeanFactor-1000}
	\end{minipage}
}
\subfloat[\sudokutest 16000]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/sudoku-sm/bestAndWorstSudoku-16000.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
    \label{sudokuMeanFactor-16000}
	\end{minipage}
}
\newline
\vspace*{0.5 cm}
\newline
\subfloat[\rmtest 11213 32]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/sm-rm/bestAndWorstRMSM-11213-32.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
	\label{sm-rmMeanFactor-11213-32}
	\end{minipage}
}
\subfloat[\rmtest 11213 64]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/sm-rm/bestAndWorstRMSM-11213-64.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
    \label{sm-rmMeanFactor-11213-64}
	\end{minipage}
}
\newline
\vspace*{0.5 cm}
\newline
\subfloat[\torustest (dist) 4096]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/distributed-torus/bestAndWorstTorus-4096.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
	\label{torusMeanFactor-4096}
	\end{minipage}
}
\subfloat[\rmtest (dist) 44497 256]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/distributed-rm/bestAndWorstRM-44497-256.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
    \label{dist-rmMeanFactor-44497}
	\end{minipage}
}
\newline
\vspace*{0.5 cm}
\newline
\subfloat[\jacobitest (dist) 3217]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/distributed-jacobi/bestAndWorstJacobi-3217.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
    \label{jacobiMeanFactor-3217}
	\end{minipage}
}
\subfloat[\jacobitest (dist) 4253]{
	\begin{minipage}[t]{.49\linewidth}%
    \centering%
    \begingroup\catcode`"=9
	\begin{tabular}{C{2cm} C{1.7cm} C{2.3cm}}
		\thead{vs.}             & \thead{mean \\ overhead} & \thead{maximum \\ normalized \\ stdDev} \\ \hline
		\csvreader[head to column names]{benchmarks/distributed-jacobi/bestAndWorstJacobi-4253.csv}{}{\vs & \meanOverhead & \maxStdDevForOverhead \\}
	\end{tabular}
	\endgroup
    \label{jacobiMeanFactor-4253}
	\end{minipage}
}
\end{table}
}

\meanOverhead