### 2017-11-29 paper/separate_plots/main.tex
wip separate plots

% This is our submission, modified from:
% the file JFP2egui.lhs
% release v1.02, 27th September 2001
%   (based on JFPguide.lhs v1.11 for LaLhs 2.09)
% Copyright (C) 2001 Cambridge University Press

\NeedsTeXFormat{LaTeX2e}

\documentclass{jfp1}

%%% Macros for the guide only %%%
%\providecommand\AMSLaTeX{AMS\,\LaTeX}
%\newcommand\eg{\emph{e.g.}\ }
%\newcommand\etc{\emph{etc.}}
%\newcommand\bcmdtab{\noindent\bgroup\tabcolsep=0pt%
%  \begin{tabular}{@{}p{10pc}@{}p{20pc}@{}}}
%\newcommand\ecmdtab{\end{tabular}\egroup}
%\newcommand\rch[1]{$\longrightarrow\rlap{$#1$}$\hspace{1em}}
%\newcommand\lra{\ensuremath{\quad\longrightarrow\quad}}

\jdate{August 2017}
\pubyear{2017}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\doi{...}

%\newtheorem{lemma}{Lemma}[section]

\input{prelude}

\title{Plots for \enquote{Arrows for Parallel Computations}}
\ifthenelse{\boolean{anonymous}}{%
\author{Submission ID xxxxxx}
}{%
%\author{Martin Braun, Phil Trinder, and Oleg Lobachev}
%\affiliation{University Bayreuth, Germany and Glasgow University, UK}
 \author[M. Braun, O. Lobachev and P. Trinder]%
        {MARTIN BRAUN\\
         University Bayreuth, 95440 Bayreuth, Germany\\
		 OLEG LOBACHEV\\
		 University Bayreuth, 95440 Bayreuth, Germany\\
		 \and\ PHIL TRINDER\\
		 Glasgow University, Glasgow, G12 8QQ, Scotland}
}% end ifthenelse

\begin{document}

\label{firstpage}

\def\SymbReg{\textsuperscript{\textregistered}}

\maketitle

\tableofcontents

\section{Plots}

\newlength{\plotwidthDist}
\setlength{\plotwidthDist}{0.6\textwidth}

\newcommand{\meanOverheadPlot}[5]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
scale only axis, width=\plotwidthDist,
xlabel=Threads,
%xtick=data,
ytick distance=#2 / 4,
xtick distance=#3,
minor tick num=9,
ylabel=Overhead (\%),
ylabel near ticks,
grid=both,
legend entries={Mean Overhead},
legend style={at={(0.99,0.99)},anchor=north east},
max space between ticks=50pt,
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
ymin=-#2,
ymax=#2,
xmin=-1,
xmax=#4]
\addplot+[mark=*,very thick,error bars/.cd,
    y dir=both,y explicit] table [x="nCores", y="overhead", y error="stdDevForOverhead", col sep=comma, mark=dots,
smooth]{#5};
\end{axis}
\end{tikzpicture}
}

\newcommand{\rmtest}{Rabin--Miller test\xspace}
\newcommand{\sudokutest}{Sudoku\xspace}
\newcommand{\jacobitest}{Jacobi sum test\xspace}
\newcommand{\torustest}{Gentleman\xspace}

\subsection{Distributed}

\subsubsection{\rmtest}

\meanOverheadPlot{Mean Overhead for distributed \\ \rmtest 44497 256}{0.02}{32}{272}{../content/benchmarks/distributed-rm/44497-256-diff.csv}

\subsubsection{\jacobitest}

\meanOverheadPlot{Mean Overhead for distributed \\ \jacobitest 4253}{0.5}{32}{272}{../content/benchmarks/distributed-jacobi/3-4253-diff.csv}
\\
\meanOverheadPlot{Mean Overhead for distributed \\ \jacobitest 3217}{0.5}{32}{272}{../content/benchmarks/distributed-jacobi/3-4253-diff.csv}

\subsubsection{\torustest}

\meanOverheadPlot{Mean Overhead for distributed \\ \torustest 4096}{0.5}{32}{272}{../content/benchmarks/distributed-torus/4096-diff.csv}

\subsection{Shared-Memory}

\subsubsection{\rmtest}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 32 vs Eden CP}{0.5}{4}{18}{../content/benchmarks/sm-rm/eden-cp-11213-32-diff.csv}
\\
\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 64 vs Eden CP}{0.5}{4}{18}{../content/benchmarks/sm-rm/eden-cp-11213-64-diff.csv}
\\
\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 32 vs GpH}{0.5}{4}{18}{../content/benchmarks/sm-rm/mult-11213-32-diff.csv}
\\
\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 64 vs GpH}{0.5}{4}{18}{../content/benchmarks/sm-rm/mult-11213-32-diff.csv}
\\
\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 32 vs Par Monad}{0.5}{4}{18}{../content/benchmarks/sm-rm/par-11213-32-diff.csv}
\\
\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 64 vs Par Monad}{0.5}{4}{18}{../content/benchmarks/sm-rm/par-11213-32-diff.csv}

    %benchmarks/sudoku-sm
    %    "eden-cp-1000-diff.csv"
    %"eden-cp-16000-diff.csv"
    %"mult-1000-diff.csv"
    %"mult-16000-diff.csv"
    %"par-1000-diff.csv"
    %"par-16000-diff.csv"

\end{document}