### 2018-01-04 paper/content/benchmarks.tex
benchmark chapter

%
\section{Performance results and discussion}
\label{sec:benchmarks}

In the following section, we describe the benchmarks of our parallel DSL and algorithmic skeletons conducted.
We start by explaining the hardware and software stack and also elaborate on which benchmarks programs were used
and also which parallel Haskell were used in which setting and why. Before we go into detail on the benchmarks we
also address the issue of poor performance gains when using hyper-threading and our reasoning behind not including
hyper-threaded cores in our benchmarks. Finally, we show that PArrows hold up in terms of performance when compared to the original parallel Haskells used as backends in this paper, starting with the shared-memory variants (GpH, |Par| Monad and Eden CP) and concluding with Eden as a distributed backend.

\newcommand{\rmtest}{Rabin--Miller test\xspace}
\newcommand{\sudokutest}{Sudoku\xspace}
\newcommand{\jacobitest}{Jacobi sum test\xspace}
\newcommand{\torustest}{Gentleman\xspace}
\newlength{\plotwidthSMP}
\setlength{\plotwidthSMP}{0.39\textwidth}
\newlength{\plotwidthDist}
\setlength{\plotwidthDist}{0.6\textwidth}

\newcommand{\benchmarkDir}{benchmarks}

\newcommand{\speedupplot}[8]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
scale only axis, width=#7,
xlabel=Threads,
%xtick=data,
%ytick=data,
xtick distance=#4,
ytick distance=#4,
ylabel=Speedup,
ylabel near ticks,
grid=major,
legend entries={linear, #2},
legend style={at={(0.01,0.99)},anchor=north west},
max space between ticks=50pt,
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
ymin=-1,
xmin=-1,
ymax=#8,
xmax=#6]
\addplot [domain=0:#3, no markers,dotted,thick]{x};
#5
\end{axis}
\end{tikzpicture}
}

\subsection{Preliminaries}

\subsubsection{Hardware and Software used}

Benchmarks were run both in a shared and in a distributed
memory setting. All benchmarks were done on the Glasgow GPG Beowulf cluster, consisting of
16 machines with 2 Intel\SymbReg~Xeon\SymbReg~E5-2640 v2 and 64 GB of DDR3 RAM each. Each processor has 8 cores and 16 (hyper-threaded) threads with a base frequency of 2 GHz and a turbo frequency of 2.50 GHz. This results in a total of 256 cores and 512 threads for the whole cluster. The operating system was Ubuntu 14.04 LTS with Kernel 3.19.0-33. Non-surprisingly, we found that hyper-threaded 32 cores do not behave in the same manner as real 16 cores (numbers here for a single machine). We disregarded the hyper-threading ability in most of the cases.

Apart from Eden, all benchmarks and libaries were compiled with Stack's\footnote{see \url{https://www.haskellstack.org/}} lts-7.1 GHC compiler which is equivalent to a standard GHC 8.0.1 with the base package in version 4.9.0.0. Stack itself was used in version 1.3.2. For GpH in its Multicore variant we used the the parallel package in version 3.2.1.0\footnote{see \url{https://hackage.haskell.org/package/parallel-3.2.1.0}}, while for the |Par| monad we used monad-par in version 0.3.4.8\footnote{see \url{https://hackage.haskell.org/package/monad-par-0.3.4.8}}. For all Eden tests, we used its GHC-Eden compiler in version 7.8.2\footnote{see \url{http://www.mathematik.uni-marburg.de/~eden/?content=build_eden_7_&navi=build}} together with OpenMPI 1.6.5\footnote{see \url{https://www.open-mpi.org/software/ompi/v1.6/}}.

Furtermore, all benchmarks were done with help of the bench\footnote{see \url{https://hackage.haskell.org/package/bench}} tool in version 1.0.2 which uses criterion (>=1.1.1.0 \&\& < 1.2)\footnote{see \url{https://hackage.haskell.org/package/criterion-1.1.1.0}} internally. All runtime data (mean runtime, max stddev, etc.) was collected with this tool if not mentioned otherwise.

We used a single node with 16 real cores as a shared memory testbed
and the whole grid with 256 real cores as a device to test our
distributed memory software.

\subsubsection{Benchmarks}

We used multiple tests that originated from different
sources. Most of them are parallel mathematical computations, initially
implemented in Eden. Table~\ref{tab:benches} summarises.

\begin{table}
\centering
\caption{The benchmarks we use in this paper.}
\label{tab:benches}
%% something was wrong with separators in table
\renewcommand{\tabcolsep}{0.5em}
\begin{tabular}{lccll}
\toprule
Name & Area & Type & Origin & Source \\
\midrule
\rmtest & Mathematics & \ensuremath{\Varid{parMap}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\jacobitest & Mathematics & \ensuremath{\Varid{workpool}\mathbin{+}\Varid{reduce}} & Eden & \citet{Lobachev2012}\\
\torustest & Mathematics & \ensuremath{\Varid{torus}} & Eden & \citet{Eden:SkeletonBookChapter02}\\
\sudokutest & Puzzle & \ensuremath{\Varid{parMap}} & \ensuremath{\Conid{Par}} Monad & \citet{par-monad}\tablefootnote{actual code from: \url{http://community.haskell.org/\~simonmar/par-tutorial.pdf and https://github.com/simonmar/parconc-examples}}\\
\bottomrule
\end{tabular}
\end{table}

\rmtest is a probabilistic primality test that iterates multiple (here: 32--256)
\enquote{subtests}. Should a subtest fail, the input is definitely not a
prime. If all $n$ subtest pass, the input is composite with the
probability of $1/4^{n}$.

Jacobi sum test or APRCL is also a primality test, that however,
guarantees the correctness of the result. It is probabilistic in the
sense that its run time is not certain. Unlike \rmtest, the subtests
of Jacobi sum test have very different durations. \citet{lobachev-phd}
discusses some optimisations of parallel APRCL. Generic parallel
implementations of \rmtest and APRCL were presented in \citet{Lobachev2012}.

\enquote{Gentleman} is a standard Eden test program, developed
for their \ensuremath{\Varid{torus}} skeleton. It implements a Gentleman's algorithm for parallel matrix
multiplication \citep{Gentleman1978}. We ported an Eden based version \citep{Eden:SkeletonBookChapter02} to PArrows.

A~parallel Sudoku solver was used by \citet{par-monad} to compare \ensuremath{\Conid{Par}} Monad
to GpH, and we ported it to PArrows.

\subsubsection{What parallel Haskells run where}

The \ensuremath{\Conid{Par}} monad and GpH -- in its multicore version \cite{Marlow2009} --  can be executed on shared memory machines only.
Although GpH is available on distributed memory
clusters, and newer distributed memory Haskells such as HdpH exist,
current support of distributed memory in PArrows is limited to
Eden. We used the MPI backend of Eden in a distributed memory
setting. However, for shared memory Eden features a ``CP'' backend
that merely copies the memory blocks between distributed heaps. In
this mode, Eden still operates in the ``nothing shared'' setting, but
is adapted better to multicore machines. We label this version of Eden
in the plots as ``Eden~CP''.

\subsubsection{Effect of hyper-threading}

In preliminary tests, the PArrows version of \rmtest on a single node of the Glasgow cluster
showed almost linear speedup on up to 16 shared-memory cores (Fig.~\ref{fig:bench-rm-sm}). The speedup
of 64-task PArrows/Eden at 16 real cores version was 13.65 giving a parallel
efficiency of 85.3\%. However, if we increased the number of
requested cores to 32 -- \ie if we use hyper-threading on 16 real
cores -- the speedup did not increase that well. It was merely 15.99
for 32 tasks with PArrows/Eden. This was worse for other backends.  As
for 64 tasks, we obtained a speedup of 16.12 with PArrows/Eden at 32
hyper-threaded cores and only 13.55 with PArrows/GpH. The parallel eficiencies were 50.4\% and 42.3\%, respectively. The Eden
version used here was Eden~CP, the \enquote{share nothing} SMP build.

In the distributed memory setting the same effect ensues. We obtain
plummeting speedup of 124.31 at 512 hyper-threaded cores, whereas it was
213.172 for 256 real cores. Apparently, hyper-threading in the Glasgow
cluster fails to execute two parallel Haskell processes with full-fledged
parallelism.

\mbcomment{fix this when benchmarks are redone}
With this in mind, the performance measurements in the Section~\ref{sec:benchmarkResults} use only real cores as the purpose of this paper is to show the performance of PArrows and not to investigate parallel performance with hyper-threading in more detail.

% rm 11213 32 32-sm speedup eden: 15.993037587283924
% -"- multi: 15.09948017762912
% -"- par: 14.909092857846693

% -"- 64 32-sm speedup eden: 16.118040224478424
% -"- multi: 13.545304115702333
% -"- par: 15.155709987503396

\subsection{Benchmark results}\label{sec:benchmarkResults}

In the following paragraphs we will go into detail on how well PArrows perform when compared to versions of the benchmark programs implemented with the original parallel Haskells that were used in the backends. We start by comparing Speedups and Overheads of for the shared memory backends and conclude with OpenMPI variants of the Eden-enabled backend as a representative of a distributed memory backend.

\subsubsection{Shared memory}

\paragraph{Speedup}
The \rmtest benchmark showed almost linear speedup for both 32 and 64 tasks, the performance is slightly better in the latter case: 13.7 at 16 cores for input $2^{11213}-1$ and 64 tasks in the best case scenario with Eden CP. The performance of the \sudokutest benchmark merely reaches a speedup of 9.19 (GpH), 8.78 (Par Monad), 8.14 (Eden CP) for 16 cores and 1000 Sudokus. In contrast to Rabin–Miller, here the |GpH| Monad backend seems to be the best of all, while Rabin–Miller profited most from Eden CP (i.e., Eden with direct memory copy) implementation of PArrows. Gentleman on shared memory has a plummeting speedup curve with GpH and |Par| Monad and logarithmically increasing speedup for Eden-based version. The latter reached a speedup of 6.56 at 16 cores.

\mbcomment{Ich hab ein bisschen mehr Daten hinzugefügt für Sudoku}

\paragraph{Overhead}
We compared the overhead of our PArrows implementations to the original parallel Haskells. We basically implemented the test programs both in PArrows and in traditional style, benchmarked both and compare the mean overhead, i.e. the \textit{relative} difference between both benchmarks with same settings. The error margin of the benchmarks, supplied by criterion package, goes into the mean overhead computation.

For the shared memory \rmtest benchmark, implemented with PArrows both with Eden CP and with |Par| Monad backends, \mbcomment{stimmt das so noch? Ich bin die Daten da nicht zu 100\% durch gegangen, aber die Aussage always finde ich etwas viel, almost always?} the zero \mbcomment{almost} always lay in the error margin of the mean overhead. This means that though we have measured some difference (against or even in favour of PArrows), it could be merely the error margin of the benchmarks and the difference might not be existent. In the following, we call the contrary case “significant”. In other words, we notice no significant slow-down through PArrows here. For GpH backend there was only a small significant overhead of $8\%$ at 16 cores with 32 tasks and input value $2^{11213}-1$.
For Sudoku the situation is slightly different. There is a minimal ($1.4\%$) significant speed improvement with PArrows Eden CP version when compared with the base Eden CP benchmark. However, with increasing number of cores the error margin reaches zero again. The |Par| Monad shows a similar development. The GpH version shows a small significant speed \textit{improvement} from PArrows of $4.2\%$.
The Gentleman multiplication with Eden CP shows a minor significant overhead of $2.6\%$ at 8 cores and an insignificant improvement at 16 cores.

We observe a low (if significant at all) overhead, induced by PArrows in the shared memory setting.

\subsubsection{Distributed Memory}

\paragraph{Speedup}
The speedup of distributed memory Rabin–Miller benchmark with PArrows and distributed Eden backend showed an almost linear speedup sans some issues in the middle because of unfortunate task distribution. As seen in Fig.~\ref{fig:rabinMillerDistSpeedup}, we reached a speedup of 213.4 with PArrrows at 256 cores (vs. 207.7 with pure Eden). We managed to measure the speedup of Jacobi sum test for sufficient large inputs only in a massively distributed setting, because of memory limitations. We improved there from 9193 seconds (128 cores) to 1649 seconds (256 cores) in our PArrows version, so we only include a graph for a run $2^{3217}-1$ in the supplementary material. This speedup at about 11 for both PArrows and Eden for more than 64 cores. There is apparently not enough work for that many cores. The Gentleman test with input 4096 had an almost linear speedup first, then plummeted between 128 and 224 cores, and recovered at 256 cores with speedup of 129.

\begin{figure}[ht]
	\centering
	\speedupplot{Speedup of distributed \rmtest \enquote{44497 256}}{PArrows Eden}{256}{32}{
	\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
	smooth]{\benchmarkDir/distributed-rm/bench-distributed.bench.skelrm-parrows-44497-256.csv};
	}{272}{\plotwidthDist}{272}
	\caption[Speedup distributed Rabin Miller]{Speedup of the distributed \rmtest benchmark with the Eden backend}
	\label{fig:rabinMillerDistSpeedup}
\end{figure}

\paragraph{Overhead}
We use our above mean overhead quality measure and the notion of significance also for distributed memory benchmarks. The mean overhead of Rabin–Miller test in the distributed memory setting ranges from $0.29\%$ to $-2.8\%$ (last value in favor of PArrows), but these values are not significant. A sole significant (by a very low margin) overhead is $0.35\%$ at 64 cores.
We measured the mean overhead for Jacobi benchmark for an input of $2^{3217}-1$ for up to 256 cores. We reached a maximum overhead of $0.32\%$ at 256 cores. In all cases it was not significant, the zero was inside the error margin.
Mean overhead for distributed Gentleman multiplication was also low. Significant values include $1.23\%$ at 64 cores, $6.4\%$ at 128 cores, and $2.4\%$ at 256 cores. It took PArrows 64.2 seconds at 256 cores to complete the benchmark.

Similar to the shared memory setting, PArrows only imply a very low penalty with distributed memory that is outweighed by their convenience and usefulness to the user.

\subsection{Discussion}

\input{bestAndWorstBenchmarks}

\mbcomment{Schau hier vll nochmal drüber, das ist etwas unsauber formuliert evtl.?}
Summing up, PArrows performed with little to no overhead. Tables \ref{tab:meanOverheadSharedMemory} and \ref{tab:meanOverHeadDistributedMemory} clarify this once more: The PArrow enabled versions trade blows with the vanilla counterparts when comparing the mean of the mean overheads. If we combine these findings with the convenience of our DSL, PArrows seem like a great improvement over the respective backends with.