### 2018-07-23 paper/separate_plots/main.tex
minor

% This is our submission, modified from:
% the file JFP2egui.lhs
% release v1.02, 27th September 2001
%   (based on JFPguide.lhs v1.11 for LaLhs 2.09)
% Copyright (C) 2001 Cambridge University Press

\NeedsTeXFormat{LaTeX2e}

\documentclass[final]{jfp1}

%%% Macros for the guide only %%%
%\providecommand\AMSLaTeX{AMS\,\LaTeX}
%\newcommand\eg{\emph{e.g.}\ }
%\newcommand\etc{\emph{etc.}}
%\newcommand\bcmdtab{\noindent\bgroup\tabcolsep=0pt%
%  \begin{tabular}{@{}p{10pc}@{}p{20pc}@{}}}
%\newcommand\ecmdtab{\end{tabular}\egroup}
%\newcommand\rch[1]{$\longrightarrow\rlap{$#1$}$\hspace{1em}}
%\newcommand\lra{\ensuremath{\quad\longrightarrow\quad}}

\jdate{January 2018}
\pubyear{2018}
\pagerange{\pageref{firstpage}--\pageref{lastpage}}
\doi{...}

%\newtheorem{lemma}{Lemma}[section]

\input{prelude}

\title{Arrows for Parallel Computation}
\subtitle{Supplementary Material}
\ifthenelse{\boolean{anonymous}}{%
\author{Submission ID xxxxxx}
}{%
%\author{Martin Braun, Phil Trinder, and Oleg Lobachev}
%\affiliation{University Bayreuth, Germany and Glasgow University, UK}
 \author[M. Braun, O. Lobachev, and P. Trinder]%
        {\textls*{MARTIN BRAUN}\\
         University Bayreuth, 95440 Bayreuth, Germany\\
		 \textls*{OLEG LOBACHEV}\\
		 University Bayreuth, 95440 Bayreuth, Germany\\
		 \and\ \textls*{PHIL TRINDER}\\
		 Glasgow University, Glasgow, G12 8QQ, Scotland}
}% end ifthenelse

% hack, but we do not need 100\% perfect rendering
\newcommand{\Conid}[1]{#1}

\begin{document}

\label{firstpage}

\def\SymbReg{\textsuperscript{\textregistered}}

\maketitle

\begin{abstract}
  This is supplementary material to our paper \enquote{Arrows for
    parallel computation}. It shows the speedup and mean overhead
  plots for all benchmarks.
\end{abstract}

\tableofcontents
%
% \newpage

\newlength{\plotwidthDist}
\setlength{\plotwidthDist}{0.6\textwidth}
\newlength{\plotwidthSMP}
% \setlength{\plotwidthSMP}{0.39\textwidth}
\setlength{\plotwidthSMP}{\plotwidthDist}

\newcommand{\meanOverheadPlot}[5]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
scale only axis, width=\plotwidthDist,
xlabel=Threads,
%xtick=data,
ytick distance=#2 / 4,
xtick distance=#3,
minor tick num=9,
ylabel=Overhead,
ylabel near ticks,
grid=both,
legend entries={Mean Overhead},
legend style={at={(0.99,0.99)},anchor=north east},
max space between ticks=50pt,
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
ymin=-#2,
ymax=#2,
xmin=-1,
xmax=#4]
\addplot+[mark=*,very thick,error bars/.cd,
    y dir=both,y explicit] table [x="nCores", y="overhead", y error="stdDevForOverhead", col sep=comma, mark=dots,
smooth]{#5};
\end{axis}
\end{tikzpicture}
}

\newcommand{\speedupplot}[8]{
\begin{tikzpicture}
\begin{axis}[title={#1},
title style={align=center},
scale only axis, width=#7,
xlabel=Threads,
%xtick=data,
%ytick=data,
xtick distance=#4,
ytick distance=#4,
ylabel=Speedup,
ylabel near ticks,
grid=major,
legend entries={linear, #2},
legend style={at={(0.01,0.99)},anchor=north west},
max space between ticks=50pt,
grid style={line width=.1pt, draw=gray!10},
major grid style={line width=.2pt,draw=gray!50},
ymin=-1,
xmin=-1,
ymax=#8,
xmax=#6]
\addplot [domain=0:#3, no markers,dotted,thick]{x};
#5
\end{axis}
\end{tikzpicture}
}

\newcommand{\rmtest}{Rabin--Miller test\xspace}
\newcommand{\sudokutest}{Sudoku\xspace}
\newcommand{\jacobitest}{Jacobi sum test\xspace}
\newcommand{\torustest}{Gentleman\xspace}

\newcommand{\benchmarkDir}{../content/benchmarks}

%%%%% hack: center all!
\centering

\section{Shared-Memory Plots}
\subsection{\rmtest}
\subsubsection{Speedup}
\speedupplot{Speedup of \rmtest \enquote{11213 32}}{Eden CP, GpH, \ensuremath{\Conid{Par}} Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-11213-32.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-11213-32.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sm-rm/bench-sm-rm.bench.skelrm-parr-par-11213-32.csv};
}{17}{\plotwidthSMP}{17}

\speedupplot{Speedup of \rmtest \enquote{11213 64}}{Eden CP, GpH, \ensuremath{\Conid{Par}} Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sm-rm/bench-sm-rm.bench.skelrm-parr-eden-cp-11213-64.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sm-rm/bench-sm-rm.bench.skelrm-parr-mult-11213-64.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sm-rm/bench-sm-rm.bench.skelrm-parr-par-11213-64.csv};
}{17}{\plotwidthSMP}{17}

\subsubsection{Overhead}
\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 32 vs Eden CP}{0.04}{4}{18}{\benchmarkDir/sm-rm/eden-cp-11213-32-diff.csv}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 64 vs Eden CP}{0.03}{4}{18}{\benchmarkDir/sm-rm/eden-cp-11213-64-diff.csv}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 32 vs GpH}{0.2}{4}{18}{\benchmarkDir/sm-rm/mult-11213-32-diff.csv}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 64 vs GpH}{0.05}{4}{18}{\benchmarkDir/sm-rm/mult-11213-64-diff.csv}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 32 vs Par Monad}{0.2}{4}{18}{\benchmarkDir/sm-rm/par-11213-32-diff.csv}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \rmtest 11213 64 vs Par Monad}{0.2}{4}{18}{\benchmarkDir/sm-rm/par-11213-64-diff.csv}

\subsection{\sudokutest}
\subsubsection{Speedup}
\speedupplot{Parallel speedup of \sudokutest \enquote{1000}}{Eden CP, GpH, \ensuremath{\Conid{Par}} Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-eden-sudoku17.1000.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-mult-sudoku17.1000.txt.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/sudoku-sm/bench-sudoku-sm.bench.parrows-sudoku-parmap-par-sudoku17.1000.txt.csv};
}{17}{\plotwidthSMP}{17}

\subsubsection{Overhead}
\meanOverheadPlot{Mean Overhead for shared-memory \\ \sudokutest 1000 vs Eden CP}{0.05}{4}{18}{\benchmarkDir/sudoku-sm/eden-cp-1000-diff.csv}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \sudokutest 1000 vs GpH}{0.05}{4}{18}{\benchmarkDir/sudoku-sm/mult-1000-diff.csv}

\meanOverheadPlot{Mean Overhead for shared-memory \\ \sudokutest 1000 vs Par Monad}{0.05}{4}{18}{\benchmarkDir/sudoku-sm/par-1000-diff.csv}

\subsection{\torustest}
\subsubsection{Speedup}
\speedupplot{Parallel speedup of \torustest \enquote{512}}{Eden CP, GpH, \ensuremath{\Conid{Par}} Monad}{16}{4}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/torus-sm/bench-torus-sm.bench.torus-matrix-parrows-eden-sm-512.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/torus-sm/bench-torus-sm.bench.torus-matrix-parrows-multicore-512.csv};
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/torus-sm/bench-torus-sm.bench.torus-matrix-parrows-parmonad-512.csv};
}{17}{\plotwidthSMP}{17}

\subsubsection{Overhead}
\meanOverheadPlot{Mean Overhead for shared-memory \\ \torustest 512 vs Eden CP}{0.08}{4}{18}{\benchmarkDir/torus-sm/eden-cp-512-diff.csv}

\section{Distributed Plots}
\subsection{\rmtest}

\subsubsection{Speedup}
\speedupplot{Speedup of distributed \rmtest \enquote{44497 256}}{PArrows Eden}{256}{32}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/distributed-rm/bench-distributed.bench.skelrm-parrows-44497-256.csv};
}{272}{\plotwidthDist}{272}

\subsubsection{Overhead}

\meanOverheadPlot{Mean Overhead for distributed \\ \rmtest 44497 256}{0.04}{32}{272}{\benchmarkDir/distributed-rm/44497-256-diff.csv}

\subsection{\jacobitest}

\subsubsection{Speedup}
\speedupplot{Speedup of distributed \jacobitest \enquote{3217}}{PArrows Eden}{256}{32}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/distributed-jacobi/bench-jacobi.bench.jacobi-parr-3-3217.csv};
}{272}{\plotwidthDist}{32}

\subsubsection{Overhead}
\meanOverheadPlot{Mean Overhead for distributed \\ \jacobitest 3217}{0.04}{32}{272}{\benchmarkDir/distributed-jacobi/3-3217-diff.csv}

\subsection{\torustest}

\subsubsection{Speedup}
\speedupplot{Speedup of distributed \torustest \enquote{4096}}{PArrows Eden}{256}{32}{
\addplot+ [very thick] table [scatter, x="nCores", y="speedup", col sep=comma, mark=none,
smooth]{\benchmarkDir/distributed-torus/bench-torus-distributed.bench.torus-matrix-parrows-eden-4096.csv};
}{272}{\plotwidthDist}{160}

\subsubsection{Overhead}
\meanOverheadPlot{Mean Overhead for distributed \\ \torustest 4096}{0.04}{32}{272}{\benchmarkDir/distributed-torus/4096-diff.csv}

\end{document}